### 2021-03-24 ch05_paper_ecs/paper_ecs.tex
Added details to changes in the representation of mixed-phase clouds in some CMIP6 models

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{ch05_paper_ecs/figs/zha.pdf}
  \caption[
    Emergent relationship ZHA for different subsets of \acs{CMIP}5 models.
  ]{
    Emergent relationship ZHA \autocite{Zhai2015} for different subsets of
    \acs{CMIP}5 models. Blue circles show the 15 \acs{CMIP}5 models used in the
    original publication (except for CESM1-CAM5). The solid blue line and blue
    shaded area show the emergent relationships evaluated on these models
    including the uncertainty range. In our study, we added eleven more
    \acs{CMIP}5 models (red circles). The corresponding emergent relationship
    that considers all available \acs{CMIP}5 models is shown in red colors.
    This relationship shows a considerably lower coefficient of determination
    ($R^2$) and higher $p$-value than the relationship using the original
    subset of \acs{CMIP}5 models. The vertical dashed line and shaded area
    correspond to the observational reference, and the horizontal dashed lines
    show the corresponding constraints on the \acf{ECS} using this observation.
    \AdaptedFrom{Schlund2020a}.
  }
  \label{fig:05:zha}
\end{figure}

\textcite{Zhai2015} focus on the variations of \acp{MBLC}, which largely
contribute to the shortwave cloud feedback and thus to the uncertainty in
modeled \ac{ECS}. Their central quantity is the response of the \ac{MBLC}
fraction to changes in the \ac{SST} in subtropical oceanic subsidence regions
for both hemispheres ($\rangeunit{20}{40}{\degree}$). On short (seasonal) and
long (centennial under a forcing) time scales, this quantity is well correlated
with \ac{ECS} among an ensemble of \acs{CMIP}3 and \acs{CMIP}5 models. Together
with observations of cloud fraction from Cloudsat/CALIPSO \autocite{Mace2009},
\ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011} and vertical velocity from
ERA-Interim \autocite{Dee2011}, the seasonal response of \ac{MBLC} fraction to
changes in \ac{SST} forms an emergent constraint on \ac{ECS}. We assess a $66
\unit{\%}$ \ac{ECS} range of $\pmrangeunit{3.35}{0.74}{K}$ for \acs{CMIP}5
($R^2 = 0.05$) and $\pmrangeunit{3.79}{0.67}{K}$ for \acs{CMIP}6 ($R^2 =
0.62$). In their original publication, \textcite{Zhai2015} find an \ac{ECS}
range of $\pmrangeunit{3.90}{0.45}{K}$ (standard deviation) for a combination
of \acs{CMIP}3 and \acs{CMIP}5 models. In terms of statistical significance,
the results of the ZHA constraints are somewhat surprising: although
\acs{CMIP}5 data (in combination with \acs{CMIP}3 data) has successfully been
used in the original publication, our approach finds that the statistical
significance of the emergent relationship is much higher in the unseen
\acs{CMIP}6 ensemble ($p < 0.0001$) than in the previously available
\acs{CMIP}5 ensemble ($p = 0.2567$). The ZHA constraint is the only emergent
constraint analyzed here that shows this extreme behavior (only one other
constraint, SHS, shows a slightly higher significance in \acs{CMIP}6; all other
constraints show lower significances in \acs{CMIP}6). The reason for the
erratic skill in \acs{CMIP}5 is the set of climate models used. For our
analysis, we use eleven additional \acs{CMIP}5 models that have not been used
in the original publication (ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m,
CCSM4, GFDL-ESM2G, GFDL-ESM2M, IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR and
MPI-ESM-P). Due to a lack of publicly available data, the model CESM1-CAM5 that
is used in the original publication is not included in our analysis. The effect
of choosing different subsets of \acs{CMIP}5 models on the emergent
relationship is illustrated in \cref{fig:05:zha}. Using the original
\acs{CMIP}5 models from the original publication gives a considerably higher
correlation ($R^2 = 0.38$) than using all available \acs{CMIP}5 models ($R^2 =
0.05$). This result shows a strong dependency of this emergent constraint on
the subset of climate models used. Nonetheless, the performance on \acs{CMIP}6
models is, surprisingly, the best of all the constraints and much better than
on either subset of \acs{CMIP}5 models.

>>>>

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{ch05_paper_ecs/figs/zha.pdf}
  \caption[
    Emergent relationship ZHA for different subsets of \acs{CMIP}5 models.
  ]{
    Emergent relationship ZHA \autocite{Zhai2015} for different subsets of
    \acs{CMIP}5 models. Blue circles show the 15 \acs{CMIP}5 models used in the
    original publication (except for \acs{CESM}1-CAM5). The solid blue line and
    blue shaded area show the emergent relationships evaluated on these models
    including the uncertainty range. In our study, we added eleven more
    \acs{CMIP}5 models (red circles). The corresponding emergent relationship
    that considers all available \acs{CMIP}5 models is shown in red colors.
    This relationship shows a considerably lower coefficient of determination
    ($R^2$) and higher $p$-value than the relationship using the original
    subset of \acs{CMIP}5 models. The vertical dashed line and shaded area
    correspond to the observational reference, and the horizontal dashed lines
    show the corresponding constraints on the \acf{ECS} using this observation.
    \AdaptedFrom{Schlund2020a}.
  }
  \label{fig:05:zha}
\end{figure}

\textcite{Zhai2015} focus on the variations of \acp{MBLC}, which largely
contribute to the shortwave cloud feedback and thus to the uncertainty in
modeled \ac{ECS}. Their central quantity is the response of the \ac{MBLC}
fraction to changes in the \ac{SST} in subtropical oceanic subsidence regions
for both hemispheres ($\rangeunit{20}{40}{\degree}$). On short (seasonal) and
long (centennial under a forcing) time scales, this quantity is well correlated
with \ac{ECS} among an ensemble of \acs{CMIP}3 and \acs{CMIP}5 models. Together
with observations of cloud fraction from Cloudsat/CALIPSO \autocite{Mace2009},
\ac{SST} from AMSRE \ac{SST} \autocite{AMSRE2011} and vertical velocity from
ERA-Interim \autocite{Dee2011}, the seasonal response of \ac{MBLC} fraction to
changes in \ac{SST} forms an emergent constraint on \ac{ECS}. We assess a $66
\unit{\%}$ \ac{ECS} range of $\pmrangeunit{3.35}{0.74}{K}$ for \acs{CMIP}5
($R^2 = 0.05$) and $\pmrangeunit{3.79}{0.67}{K}$ for \acs{CMIP}6 ($R^2 =
0.62$). In their original publication, \textcite{Zhai2015} find an \ac{ECS}
range of $\pmrangeunit{3.90}{0.45}{K}$ (standard deviation) for a combination
of \acs{CMIP}3 and \acs{CMIP}5 models. In terms of statistical significance,
the results of the ZHA constraints are somewhat surprising: although
\acs{CMIP}5 data (in combination with \acs{CMIP}3 data) has successfully been
used in the original publication, our approach finds that the statistical
significance of the emergent relationship is much higher in the unseen
\acs{CMIP}6 ensemble ($p < 0.0001$) than in the previously available
\acs{CMIP}5 ensemble ($p = 0.2567$). The ZHA constraint is the only emergent
constraint analyzed here that shows this extreme behavior (only one other
constraint, SHS, shows a slightly higher significance in \acs{CMIP}6; all other
constraints show lower significances in \acs{CMIP}6). The reason for the
erratic skill in \acs{CMIP}5 is the set of climate models used. For our
analysis, we use eleven additional \acs{CMIP}5 models that have not been used
in the original publication (ACCESS1-0, ACCESS1-3, bcc-csm1-1, bcc-csm1-1-m,
CCSM4, GFDL-ESM2G, GFDL-ESM2M, IPSL-CM5A-MR, IPSL-CM5B-LR, MPI-ESM-MR and
MPI-ESM-P). Due to a lack of publicly available data, the model
\acs{CESM}1-CAM5 that is used in the original publication is not included in
our analysis. The effect of choosing different subsets of \acs{CMIP}5 models on
the emergent relationship is illustrated in \cref{fig:05:zha}. Using the
original \acs{CMIP}5 models from the original publication gives a considerably
higher correlation ($R^2 = 0.38$) than using all available \acs{CMIP}5 models
($R^2 = 0.05$). This result shows a strong dependency of this emergent
constraint on the subset of climate models used. Nonetheless, the performance
on \acs{CMIP}6 models is, surprisingly, the best of all the constraints and
much better than on either subset of \acs{CMIP}5 models.

### 2021-03-24 ch07_conclusion/conclusion.tex
Added details to changes in the representation of mixed-phase clouds in some CMIP6 models

In the first study of this thesis, the climate sensitivity metrics \ac{ECS} and
\ac{TCR} are evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6.
This work, which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and
already published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found,
which is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds over the Southern Ocean. This change leads
to an improved simulation of the shortwave \ac{CRE} in these models when
compared to observations. However, this change also substantially reduces the
strong negative shortwave cloud feedback over the Southern Ocean that is
present in previous \ac{CMIP} generations and that results from a cloud phase
change from ice clouds in the present-day to liquid clouds in the future. In
the affected \acs{CMIP}6 models, this cloud phase change due to warming is
reduced since these models simulate less cloud ice over the present-day
Southern Ocean than their predecessor versions with no supercooled cloud
liquid.

>>>>

In the first study of this thesis, the climate sensitivity metrics \ac{ECS} and
\ac{TCR} are evaluated for the latest generation of \acp{ESM} from \acs{CMIP}6.
This work, which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and
already published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$ is found,
which is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by a third of the \acs{CMIP}6 models, with
many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. One possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is the addition of prognostic
aerosol schemes that include aerosol-cloud interactions which could have
produced overly large negative radiative forcing in some models. To correctly
reproduce the observed historical temperature trend, a stronger model response
to \ac{GHG} forcing is required, resulting in a higher climate sensitivity. A
further reason for the high climate sensitivity in \acs{CMIP}6 is a change in
the microphysical representation of mixed-phase clouds over the Southern Ocean
in some models. This change was mainly implemented to address a long-standing
cloud phase bias over the Southern Ocean in the climate models, which has been
successfully reduced. Moreover, the \acs{CMIP}6 \ac{MMM} shows an improved
simulation of the shortwave \ac{CRE} when compared to observations. However,
the change in the representation of mixed-phase clouds also substantially
reduces the strong negative shortwave cloud feedback over the Southern Ocean
that is present in previous \ac{CMIP} generations and that results from a cloud
phase change from ice clouds in the present-day to liquid clouds in the future.
In the affected \acs{CMIP}6 models, this cloud phase change due to warming is
reduced since these models simulate less cloud ice over the present-day
Southern Ocean than their predecessor versions with no supercooled cloud
liquid.

### 2021-03-24 end_abbreviations/abbreviations.tex
Restructured thesis and removed unnecessary parts
Added list of figures and list of tables
Added details on carbon cycle models to chapter 02
Added details to changes in the representation of mixed-phase clouds in some CMIP6 models

%=============================================================================%
%                                Dissertation                                 %
%                               Manuel Schlund                                %
%                                  (c) 2020                                   %
%=============================================================================%
%                                  Acronyms                                   %
%=============================================================================%

\chapter{List of Acronyms}

\begingroup
\setstretch{0.5}
\begin{acronym}
  \acro{1BGC}[1\%BGC]{biogeochemically-coupled simulation where the %
    atmospheric CO\textsubscript{2} concentration increases by $1 \unit{\%}$ %
    per year}
  \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
  \acro{AI}{artificial intelligence}
  \acro{AOGCM}{atmosphere-ocean general circulation model}
  \acro{AR}{Assessment Report}
  \acro{C4MIP}{Coupled Climate-Carbon Cycle Model Intercomparison Project}
  \acro{CESM}{Community Earth System Model}
  \acro{CFMIP}{Cloud Feedback Model Intercomparison Project}
  \acro{CMIP}{Coupled Model Intercomparison Project}
  \acro{CH4}[CH\textsubscript{4}]{methane}
  \acro{CMOR}{Climate Model Output Rewriter}
  \acro{CO}{carbon monoxide}
  \acro{CO2}[CO\textsubscript{2}]{carbon dioxide}
  \acro{CRE}{cloud radiative effect}
  \acro{CV}{cross-validation}
  \acro{DECK}{Diagnostic, Evaluation, and Characterisation of Klima}
  \acro{DJF}{December-January-February}
  \acro{ECS}{effective climate sensitivity}
  \acro{ENSO}{El Niño-Southern Oscillation}
  \acro{ESGF}{Earth System Grid Federation}
  \acro{ESM}{Earth system model}
  \acro{ESMValCore}{core functionalities of the ESMValTool}
  \acro{ESMValTool}{Earth System Model Evaluation Tool}
  \acro{GBRT}{gradient boosted regression tree}
  \acro{GHG}{greenhouse gas}
  \acrodefplural{GHG}{greenhouse gases}
  \acro{GPP}{gross primary production}
  \acro{GSAT}{global mean near-surface air temperature}
  \acro{GtC}{gigatonnes of carbon}
  \acro{IAM}{integrated assessment model}
  \acro{IPCC}{Intergovernmental Panel on Climate Change}
  \acro{ITCZ}{Intertropical Convergence Zone}
  \acro{JJA}{June-July-August}
  \acro{KUM}{Cape Kumukahi, Hawaii}
  \acro{LAI}{leaf area index}
  \acro{LASSO}{least absolute shrinkage and selection operator}
  \acro{LIME}{local interpretable model-agnostic explanations}
  \acro{LongRunMIP}{Long Run Model Intercomparison Project}
  \acro{LTMI}{lower-tropospheric mixing index}
  \acro{MBLC}{marine boundary layer cloud}
  \acro{MDER}{multiple diagnostic ensemble regression}
  \acro{MIP}{Model Intercomparison Project}
  \acro{ML}{machine learning}
  \acro{MMM}{multi-model mean}
  \acro{N2O}[N\textsubscript{2}O]{nitrous oxide}
  \acro{NAO}{North Atlantic Oscillation}
  \acro{NCAR}{National Center for Atmospheric Research}
  \acro{netCDF}{Network Common Data Form}
  \acro{PDF}{probability density function}
  \acro{ppm}{parts per million}
  \acro{PR}{precipitation}
  \acro{RCP}{Representative Concentration Pathway}
  \acro{RMSE}{root mean square error}
  \acro{RMSEP}{root mean square error of prediction}
  \acro{RSDS}{incoming solar radiation at surface}
  \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
  \acro{SPE}{standard prediction error}
  \acro{SSP}{Shared Socioeconomic Pathway}
  \acro{SST}{sea surface temperature}
  \acro{T}{near-surface air temperature}
  \acro{TCR}{transient climate response}
  \acro{TLC}{tropical low-level cloud}
  \acro{TOA}{top of the atmosphere}
  \acro{WCRP}{World Climate Research Programme}
  \acro{WGCM}{Working Group on Coupled Modelling}
\end{acronym}
\endgroup

>>>>

%=============================================================================%
%                            Doctoral Dissertation                            %
%                               Manuel Schlund                                %
%                                  (c) 2021                                   %
%=============================================================================%
%                                Abbreviations                                %
%=============================================================================%

\chapter{List of Abbreviations}

\begingroup
\setstretch{0.5}
\begin{acronym}
  \acro{1BGC}[1\%BGC]{biogeochemically-coupled simulation where the %
    atmospheric \ch{CO2} concentration increases by $1 \unit{\%}$ per year}
  \acro{AerChemMIP}{Aerosol Chemistry Model Intercomparison Project}
  \acro{AI}{artificial intelligence}
  \acro{AOGCM}{atmosphere-ocean general circulation model}
  \acro{AR}{Assessment Report}
  \acro{C4MIP}{Coupled Climate-Carbon Cycle Model Intercomparison Project}
  \acro{CESM}{Community Earth System Model}
  \acro{CFMIP}{Cloud Feedback Model Intercomparison Project}
  \acro{CH4}[\ch{CH4}]{methane}
  \acro{CMIP}{Coupled Model Intercomparison Project}
  \acro{CMOR}{Climate Model Output Rewriter}
  \acro{CO}{carbon monoxide}
  \acro{CO2}[\ch{CO2}]{carbon dioxide}
  \acro{CRE}{cloud radiative effect}
  \acro{CV}{cross-validation}
  \acro{DECK}{Diagnostic, Evaluation, and Characterisation of Klima}
  \acro{DJF}{December-January-February}
  \acro{ECS}{effective climate sensitivity}
  \acro{ENSO}{El Niño-Southern Oscillation}
  \acro{ESGF}{Earth System Grid Federation}
  \acro{ESM}{Earth system model}
  \acro{ESMValCore}{core functionalities of the ESMValTool}
  \acro{ESMValTool}{Earth System Model Evaluation Tool}
  \acro{GBRT}{gradient boosted regression tree}
  \acro{GHG}{greenhouse gas}
  \acrodefplural{GHG}{greenhouse gases}
  \acro{GPP}{gross primary production}
  \acro{GSAT}{global mean near-surface air temperature}
  \acro{GtC}{gigatonnes of carbon}
  \acro{HadGEM}{Hadley Centre Global Environmental Model}
  \acro{IAM}{integrated assessment model}
  \acro{IPCC}{Intergovernmental Panel on Climate Change}
  \acro{ITCZ}{Intertropical Convergence Zone}
  \acro{JJA}{June-July-August}
  \acro{KUM}{Cape Kumukahi, Hawaii}
  \acro{LAI}{leaf area index}
  \acro{LASSO}{least absolute shrinkage and selection operator}
  \acro{LIME}{local interpretable model-agnostic explanations}
  \acro{LongRunMIP}{Long Run Model Intercomparison Project}
  \acro{LTMI}{lower-tropospheric mixing index}
  \acro{MBLC}{marine boundary layer cloud}
  \acro{MDER}{multiple diagnostic ensemble regression}
  \acro{MIP}{Model Intercomparison Project}
  \acro{ML}{machine learning}
  \acro{MMM}{multi-model mean}
  \acro{N2O}[\ch{N2O}]{nitrous oxide}
  \acro{NAO}{North Atlantic Oscillation}
  \acro{NBP}{net biome production}
  \acro{NCAR}{National Center for Atmospheric Research}
  \acro{NEP}{net ecosystem production}
  \acro{netCDF}{Network Common Data Form}
  \acro{NPP}{net primary production}
  \acro{PDF}{probability density function}
  \acro{ppm}{parts per million}
  \acro{PR}{precipitation}
  \acro{RCP}{Representative Concentration Pathway}
  \acro{RMSE}{root mean square error}
  \acro{RMSEP}{root mean square error of prediction}
  \acro{RSDS}{incoming solar radiation at surface}
  \acro{RuBisCO}{ribulose-1,5-bisphosphate carboxylase/oxygenase}
  \acro{ScenarioMIP}{Scenario Model Intercomparison Project}
  \acro{SPE}{standard prediction error}
  \acro{SSP}{Shared Socioeconomic Pathway}
  \acro{SST}{sea surface temperature}
  \acro{T}{near-surface air temperature}
  \acro{TCR}{transient climate response}
  \acro{TLC}{tropical low-level cloud}
  \acro{TOA}{top of the atmosphere}
  \acro{WCRP}{World Climate Research Programme}
  \acro{WGCM}{Working Group on Coupled Modelling}
\end{acronym}
\endgroup
